Computing in the Field of Rationals

### 2012-06-26 rational_ext.tex
_

The rule {{(eta*)}} is different from other rules encountered before
due to the {{forall v:b}} quantification and is to be understood as
follows: when the machine encounters an {{eta*}} reduction it
non-deterministically replicates into several machine states, each of
which is seeded by a unique value {{v}} that inhabits the type {{b}}
that indexes {{eta* : 1 <-> 1/b * b}}. As an example at the type
{{bool}}, we would have two possible machine states:

@@

Logical reversibility of {{PiEE*}} implies that {{c^{dagger} }}
denotes the inverse relation that {{c}} denotes, i.e.
{{c^^^={|(v',v)~|~(v,v') elem c|} }}.  Operationally, for any
execution path from {{v}} to {{v'}} in {{c}}, {{c{dagger} }} has an
execution path from {{v'}} to {{v}}.

@@

As with {{Pi}}, {{PiEE*}} is a dagger symmetric monoidal category with
two monoidal structures. In this section we establish that it is
``locally compact closed'' in the sense that for non-zero objects,
namely types where {{[^ b ^]=0}}, {{PiEE*}} satisfies the conditions
for compact closed categories.

>>>>

The rule {{(eta*)}} is different from other rules encountered before
due to the {{forall v:b}} quantification and is to be understood as
follows: when the machine encounters an {{eta*}} reduction it
non-deterministically replicates into several machine states or
worlds, each of which is seeded by a unique value {{v}} that inhabits
the type {{b}} that indexes {{eta* : 1 <-> 1/b * b}}. Each world has
an independant existence. As an example at the type {{bool}}, we would
have two possible machine states:

@@

Logical reversibility of {{PiEE*}} implies that {{c^{dagger} }}
denotes the inverse relation that {{c}} denotes, i.e.
{{c^^^={|(v',v)~|~(v,v') elem c|} }}.  Operationally, for any
execution path from {{v}} to {{v'}} in {{c}}, {{c{dagger} }} has an
execution path from {{v'}} to {{v}}.  Intuitively, logical
reversibility holds because for every {{eta*}} reduction in a world
that transforms {{()}} into {{(1/v,v)}}, on reverse execution the
adjoint {{eps*}} will transform {{(1/v,v)}} back to {{()}}.

@@

Like {{Pi}}, {{PiEE*}} is a commutative bimonoidal catgeory.
Symmetric monoidal categories where every object had a dual object and
where dualizing preserves the monodal tensor are referred to as
compact closed categories. Since divison is defined only for types of
non-zero size, not all objects have dual objects.  In this section
establish that {{PiEE*}} is ``locally compact closed'' in the sense
that for non-zero objects, namely types where {{[^ b ^]=0}}, it
satisfies the conditions for compact closed categories.
